---
title: 'STAT 6910: HW 7'
author: "David Angeles"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    df_print: paged
fontsize: 12pt
---

```{r, echo=FALSE}
library(pwr)
library("emmeans")
library("dae")
paper.towel.strength <- read.table("~/Desktop/Stats 6910/HW_4_and_5/paper.towel.strength.txt",
                                   header = TRUE)
options(contrasts = c("contr.sum", "contr.poly"))
```


## Problem 8

Burt Beiter, Doug Fairchild, Leo Russo, and Jim Wirtley, in 1990, ran an experiment to compare the relative strengths of two similarly priced brands of paper towel under varying levels of moisture saturation and liquid type. The treatment factors were “amount of liquid” (factor $A$, with levels 5 and 10 drops coded 1 and 2), “brand of towel” (factor $B$, with levels coded 1 and 2), and “type of liquid” (factor $C$, with levels “beer” and “water” coded 1 and 2). A 2 × 2 × 2 factorial experiment with $r = 3$ was run in a completely randomized design. 


__(a)__ The experimenters assumed only factors A and B would interact. Specify the corresponding model.

__Solution__

Let $A, B$ and $C$ be as described with the associated levels $a= 2$, $b = 2$, and $c= 2$ respectively with $r=3$. Then we have that the corresponding model is


$$
Y_{ijk t} = \mu + \alpha_i + \beta_j + \gamma_k + (\alpha \beta)_{ij} + \epsilon_{ijk t}
$$
where $i = 1, 2$, $j =1, 2$, $k= 1,2$, and $t=1,2,3$.

__(b)__ Assume there is only one contrast of primary interest: the one comparing brands of paper towels.

__Solution__
Assuming that the contrast of primary interest is the one that compares the brands of paper towels, then the following contracts can be used:

$$
\bar{\tau}_{\bullet 1} - \bar{\tau}_{\bullet 2} = \{\beta_1 + (\bar{\alpha \beta})_{\bullet 1} \} - \{\beta_2 + (\bar{\alpha \beta})_{\bullet 2} \}).
$$


__(c)__ Use residual plots to evaluate the adequacy of the model specified in part (a).

For part (c), comment on the potential for interactions using the graphing tool in the dae R package; you can just use these plots (and/or two-factor interaction plots) to assess whether the model for the mean is reasonable (rather than producing further residual plots to check that assumption).

__Solution__

```{r}
paper.towel.strength <- within(paper.towel.strength,{
  A = factor(A); B = factor(B); C = factor(C); ABC = factor(ABC)})

attach(paper.towel.strength)
interaction.ABC.plot(strength, A, B, C, data = paper.towel.strength)
interaction.plot(A,C , strength , col = c(1,2,3), lwd = 2,
                 xlab = "Amount of Liquid", ylab = "Strength", 
                 trace.label = "Type of Liquid")
interaction.plot(B,C , strength , col = c(1,2,3), lwd = 2, 
                 xlab = "Towel Brand",ylab = "Strength", 
                 trace.label = "Type of Liquid")
interaction.plot(A,B , strength , col = c(1,2,3), lwd = 2, 
                 xlab = " Amount of Liquid",ylab = "Strength", 
                 trace.label = "Towel Brand")
```


The interaction plot of $A$ and $C$, Amounts of Liquid vs. Strength with respect to Type of Liquid, and the interaction plot of $B$ and $C$, Towel Brand vs. Strength with respect to Type of Liquid, both appear to have parallel lines. Furthermore, the three way interaction plot appears to have parallel lines as well but a minor downward shift of the factor $B$ when changing from level 1 to level 2 of factor $C$. 
However, it still might not be cause for concern since the shift is fairly small. 
These two results would indicate that factors $A$ and $C$, and $B$ and $C$ don't interact. Also, there probably isn't any three-way interaction either.
From the last interaction plot which shows the interactions between $A$ and $B$, we can see that the lines appear to be very parallel which would suggest that the assumptions that $A$ and $B$ interact is likely to be false and therefore the model in part (a) might not be the best model to represent the data. 


(d) Provide an analysis of variance table for this experiment, test the various effects, show plots of significant main effects and interactions, and draw conclusions.

For part (d), proceed with the typical ANOVA, without variable transformation, etc., regardless of your opinion on the residual plots in (c). Test each effect in the ANOVA table at 0.01 (so you maintain a FWER of 0.04). You do not need to produce plots of significant effects.

```{r}
model_AB <- aov(strength ~ A + B + C + A:B, data= paper.towel.strength)
anova(model_AB)
```

From the ANOVA table above we get a $p$-value of $.8456 > .01$ for the interactions between factor $A$ and $B$. So we can see that our data shows negligible interactions between Amount of Liquid and Towel Brand. The effects of Type of Liquid averaged over Amount of Liquid and Towel Brand appears to have some effect on mean strength since we obtain a p-value of $.008< .01$. The effects of Towel Brand averaged over Amount of Liquid and Type of Liquid appears to have some effect on mean strength since we obtain a p-value of $.0000028< .01$. Lastly, the Amount of Liquid averaged over Towel Brand and Type of Liquid does not seem to have an effect on mean strength since we obtain a p-value of $.068 > .01$.

(e) Construct confidence intervals for each of the treatment contrasts 
that you listed in part (b), using an appropriate method of multiple comparisons. 
Discuss the results.

For part (e), use a confidence level of 99% for the one primary contrast of interest


```{r}
em.B <- emmeans(model_AB, specs = ~B)

summary(contrast(em.B, method = "tukey"), infer = c(TRUE,TRUE), level=.99)

```


With 99% confidence, we believe that the true difference between Towel Brand 1 and Towel Brand 2 averaged over Amount of Liquid and Type of Liquid is between 471.87 and 1203.31.



