---
title: 'STAT 6910: HW 6'
author: "David Angeles"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    df_print: paged
fontsize: 12pt
---

```{r, echo=FALSE}
library(pwr)
library("emmeans")
```


## Problem 1

Under what circumstances should the two-way main effects model (6.2.3) be used rather than the two-way complete model (6.2.2)? [1 sentence may suffice.] Discuss the interpretation of main effects in each model. [You can focus on comparing the interpretations of, e.g., $\alpha_1 - \alpha_2$ vs. $\alpha_1^\star - \alpha_2^\star$ in these models.]


## Problem 2


Verify that $(\tau_{ij} - \bar{\tau}_{i\bullet} - \bar{\tau}_{\bullet j} + \bar{\tau}_{\bullet \bullet})$ is an interaction contrast for the two-way complete model. Write down the list of contrast coefficients in terms of the  $\tau_{ij}$’s when factor $A$ has $a = 3$ levels and factor $B$ has $b=4$ levels. [For the second part (“Write down the list of contrast coefficients...”) set $i=1$ and $j = 3$.]



## Problem 7 part (a)


The data shown in Table 6.22 are a subset of the data given by Anderson and McLean (1974) and show the strength of a weld in a steel bar. Two factors of interest were gage bar setting (the distance the weld die travels during the automatic weld cycle) and time of welding (total time of the automatic weld cycle). Assume that the levels of both factors were selected to be equally spaced.


```{r}
weld.data = read.table("~/Desktop/Stats 6910/HW_4_and_5/weld.strength.txt", 
                            header = TRUE)
```

(a) Using the cell-means model (6.2.1) for these data, test the hypothesis that there is no difference in the effects of the treatment combinations on weld strength against the alternative hypothesis that at least two treatment combinations have different effects.

## Problem 8

For the experiment described in Exercise 7, use the two-way complete model instead of the equivalent cell means model.

(a) Test the hypothesis of no interaction between gage bar setting and time of weld and state your conclusion.

(b) Draw an interaction plot for the two factors Gage bar setting and Time of welding. Does your interaction plot support the conclusion of your hypothesis test? Explain.

(c) In view of your answer to part (b), is it sensible to investigate the differences between the effects of gage bar setting? Why or why not? Indicate on your plot what would be compared.


## Problem 21 

The experiment was run in order to examine the amount of time taken to boil a given amount of water on the four different burners of her stove, and with 0, 2, 4, or 6 teaspoons of salt added to the water. Thus the experiment had two treatment factors with four levels each. The experimenter ran the experiment as a completely randomized design by taking $r = 3$ observations on each of the 16 treatment combinations in a random order. The data are shown in Table 6.26. The experimenter believed that there would be no interaction between the two factors.

```{r}
boiling.data = read.table("~/Desktop/Stats 6910/HW_4_and_5/water.boiling.txt", 
                            header = TRUE)
```

(a) Check the assumptions on the two-way main-effects model.

(b) Calculate a 99% set of Tukey confidence intervals for pairwise differences between the levels of salt, and calculate separately a 99% set of intervals for pairwise differences between the levels of burner.






